Distinguishing Spin-Aligned and Isotropic Black Hole Populations With Gravitational Waves

### 2017-06-02 paper/AlignedIsotropicGW.tex
Close #46.

\begin{abstract}
  The first direct detections of gravitational
  waves\cite{2016PhRvL.116f1102A,2016PhRvL.116x1103A,O1-BBH} from
  merging binary black holes open a unique window into the binary
  black hole formation environment.  One promising signature of the
  formation environment is the angular distribution of the black hole
  spins; systems formed through dynamical interactions among
  already-compact objects are expected to have isotropic spin
  orientations\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016,2016ApJ...832L...2R}
  whereas binaries formed from pairs of stars born together are more
  likely to have spins preferentially aligned with the binary orbital
  angular momentum as a consequence of their joint evolution toward a
  binary black hole
  system\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
  We consider existing gravitational wave measurements of the binary
  effective spin, the best-measured combination of spin
  parameters\cite{2016PhRvL.116x1102A,O1-BBH}, in the three likely
  binary black hole detections GW150914, LVT151012, and GW151226.  If
  binary black hole spin magnitudes extend to high values we show that
  the data already exhibit a $\OOneSigmaIsoAligned\sigma$
  ($\OOneOddsIsoAligned$ odds ratio\footnote{An odds ratio of $r$ with
    $r \ll 1$ is equivalent to $x \sigma$ with
    $x = \Phi^{-1}\left( 1 - r/2 \right)$, where $\Phi$ is the unit
    normal CDF.})  preference for an isotropic angular distribution.
  By considering the effect of 10 additional
  detections\cite{2016ApJ...833L...1A} of systems, we show that if all
  observations come from a single population such an augmented data
  set would typically enable at least a
  $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
  ratio) distinction between the isotropic and aligned models for the
  assumed spin magnitude distributions, and in most cases better than
  $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).  The existing
  preference for either an isotropic spin distribution or low spin
  magnitudes for the observed systems will be confirmed (or
  overturned) confidently in the near future by subsequent
  observations.
\end{abstract}

>>>>

\begin{abstract}
  The first direct detections of gravitational
  waves\cite{2016PhRvL.116f1102A,2016PhRvL.116x1103A,O1-BBH} from
  merging binary black holes open a unique window into the binary
  black hole formation environment.  One promising signature of the
  formation environment is the angular distribution of the black hole
  spins; systems formed through dynamical interactions among
  already-compact objects are expected to have isotropic spin
  orientations\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016,2016ApJ...832L...2R}
  whereas binaries formed from pairs of stars born together are more
  likely to have spins preferentially aligned with the binary orbital
  angular momentum as a consequence of their joint evolution toward a
  binary black hole
  system\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
  We consider existing gravitational wave measurements of the binary
  effective spin, the best-measured combination of spin
  parameters\cite{2016PhRvL.116x1102A,O1-BBH}, in the three likely
  binary black hole detections GW150914, LVT151012, and GW151226.  If
  binary black hole spin magnitudes extend to high values we show that
  the data already exhibit a $\OOneSigmaIsoAligned\sigma$
  ($\OOneOddsIsoAligned$ odds ratio\footnote{An odds ratio of $r$ with
    $r \ll 1$ is equivalent to $x \sigma$ with
    $x = \Phi^{-1}\left( 1 - r/2 \right)$, where $\Phi$ is the unit
    normal CDF.})  preference for an isotropic angular distribution
  over an aligned one.  By considering the effect of 10 additional
  detections\cite{2016ApJ...833L...1A} of systems, we show that if all
  observations come from a single population such an augmented data
  set would typically enable at least a
  $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
  ratio) distinction, and in most cases better than $5\sigma$
  ($2.9 \times 10^{-7}$ odds ratio).  The existing preference for
  either an isotropic spin distribution or low spin magnitudes for the
  observed systems will be confirmed (or overturned) confidently in
  the near future by subsequent observations.
\end{abstract}
